Multi-Domain Adaptation in Neural Machine Translation with Dynamic Sampling Strategies

### 2021-05-11 2021/emnlp/auto_cl_mdmt.tex
small changes in intro

This problem is often approached by \emph{static} instance selection or re-weighting strategies, where the available training data is used in proportion to its relevance for the chosen test conditions \citep{Moore10selection,Axelrod11domain}. Finding the optimal balance of training data is however a challenging task due, for instance, to the similarity between domains/languages, but also due to regularization effects of out-of-domain data; it may also be suboptimal, as some target domains or languages might be easier to train than others. For example, improving the performance of the MT system in one domain will often hurt that of another \citep{Vanderwees17dynamic, Britz17mixing}, and improving model generalization across all domains \citep{koehn18findings} may not promise optimization of a particular domain.

Several recent proposals \citep{platanios19competence,Zhang19curriculum,Kumar19reinforcement,Wang20learning-multi,Graves17automated,Wang20balancing,Wang20learning-multi,Vanderwees17dynamic} have explored ways to instead consider \emph{dynamic} data selection and sampling strategies that surpass static strategies. Furthermore \citet{Zhang19curriculum,Vanderwees17dynamic} construct a static curriculum while \citet{platanios19competence,Kumar19reinforcement,Graves17automated,Wang20balancing,Wang20learning-multi} build curricula automatically adapted to the training.

The major contribution of our work consists of 2 parts. First, we propose a novel framework that covers the domain adaptation problem and the multi-domain adaptation problem. Secondly, we propose a novel method of building an automated curriculum for training the Multi-Domain NMT model. We evaluate our method in various single-domain adaptation and multi-domain adaptation settings. We compare our method with several contrasts including the work of \citet{Zhang19curriculum} and \citet{Wang20balancing}. From now on, we name our method MDAC which is brief for Multi-Domain Automated Curriculum. Based on experimental results obtained on a diverse set of domains, our main conclusions are that (a) using MDAC often yields overall performance that is as good as the standard fine-tuning strategy for domain adaptation; (b) MDAC can effectively handle a variety of test target distributions.

>>>>

This problem is often approached by \emph{static} instance selection or re-weighting strategies, where the available training data is used in proportion to its relevance for the chosen test conditions \citep{Moore10selection,Axelrod11domain}. Finding the optimal balance of training data is however a challenging task due, for instance, to the similarity between domains/languages, but also due to regularization effects of out-of-domain data. It may also be suboptimal, as some target domains or languages might be easier to train than others. For example, improving the performance of the MT system in one domain will often hurt that of another \citep{Vanderwees17dynamic, Britz17mixing}, and improving model generalization across all domains \citep{koehn18findings} may not promise optimization of a particular domain.

Several recent proposals
% \citep{Vanderwees17dynamic,Graves17automated,platanios19competence,Zhang19curriculum,Kumar19reinforcement,Wang20learning-multi,Wang20balancing,Wang20learning-multi}
have explored ways to instead consider \emph{dynamic} data selection and sampling strategies that surpass static strategies. Notably,  \citet{Vanderwees17dynamic,Zhang19curriculum} construct a static curriculum, while\citet{Graves17automated,platanios19competence,Kumar19reinforcement,Wang20balancing,Wang20learning-multi} build curricula that automatically adapt to the training data.

In this paper, we contribute to this line of research in several ways. First, we propose a novel framework (\emph{Multi-Domain Automated Curriculum}, MDAC for short) that simultaneously accounts for the domain adaptation and the multi-domain adaptation problems. Second, we propose a novel method of building an automated curriculum for training Multi-Domain NMT models. We evaluate our method in various single-domain and multi-domain adaptation settings. We compare our method with several contrasts including the work of \citet{Zhang19curriculum} and \citet{Wang20balancing}.
% From now on, we name our method MDAC which is brief for .
Based on these experimental results, our main conclusions are that (a) using MDAC often yields overall performance that is as good as the standard finetuning strategy for domain adaptation; (b) MDAC can effectively handle a variety of test situations, from targeting one single domain to the full multi-domain scenario. % distributions
